---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# konfound

The goal of konfound is to carry out sensitivity analysis based on Rubin's causal model.

### Installation

You can install konfound from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("jrosen48/rsensitivity")
```

### Use 

`pkonfound()`, for published studies, calculates (1) how much bias there must be in an estimate to invalidate/sustain an inference; (2) the impact of an omitted variable necessary to invalidate/sustain an inference for a regression coefficient:

```{r}
library(konfound)
pkonfound(2, .4, 100, 3)
```

`konfound()` calculates the same for models fit in R. For example, here is the output from a linear model fit with `lm()` using the built-in dataset `mtcars`:

```{r}
m1 <- lm(mpg ~ wt + hp, data = mtcars)

summary(m1)
```

Sensitivity analysis for the effect for `wt` on `mpg` can be carried out as follows, specifying the fitted model object:

```{r}
konfound(m1, wt)
```