---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# konfound

The goal of konfound is to carry out sensitivity analysis as described in Frank, Maroulis, Duong, and Kelcey (2013) based on Rubin's (1974) causal model. 

# Installation

You can install konfound from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("jrosen48/rsensitivity")
```

```{r}
devtools::load_all(".")
```

# Use of konfound

### pkonfound() for published studies

`pkonfound()`, for published studies, calculates (1) how much bias there must be in an estimate to invalidate/sustain an inference; (2) the impact of an omitted variable necessary to invalidate/sustain an inference for a regression coefficient:

```{r, eval = F}
library(konfound)
```

```{r}
pkonfound(2, .4, 100, 3, to_return = "df")
pkonfound(2, .4, 100, 3, to_return = "print")
pkonfound(2, .4, 100, 3, to_return = "plot")
```

### konfound() for models fit in R

`konfound()` calculates the same for models fit in R. For example, here is the output from a linear model fit with `lm()` using the built-in dataset `mtcars`:

```{r}
m1 <- lm(mpg ~ wt + hp, data = mtcars)
summary(m1)
```

Sensitivity analysis for the effect for `wt` on `mpg` can be carried out as follows, specifying the fitted model object:

```{r}
konfound(m1, wt)
```

### mkonfound for meta-analyses including sensitivity analysis

```{r, eval = T}
library(dplyr, warn.conflicts = FALSE)

df <- tribble(
  ~unstd_beta, ~std_err, ~n_obs, ~n_covs,
  2,           .3,       70,     3,
  10,          2.9,      405,    4,
  1.7,         1.5,      200,    1
)

mkonfound(df)
```

# Shiny Version for published studies 

A shiny version for sensitivity analysis for published studies is available [here](https://jmichaelrosenberg.shinyapps.io/shinykonfound/).